---
title: 'MTA Shutdown Analysis'
description: |
  An automated notebook built to investigate structural patterns resulting from New York City & State transportation policy.
author:
  - name: Jason Bixon 
    url: https://jbixon13.wixsite.wix.com/website
date: "`r Sys.Date()`"
output: radix::radix_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE, echo=FALSE}
library(tidyverse)
library(lubridate)
library(plotly)
library(leaflet)
library(forecast)
library(scales)
```

```{r message=FALSE, echo=FALSE}
# read in datasets 
NYCHA <- read_csv('nycha-resident-data-book-summary.csv')                                                                # NYC Housing Authority data 
MTA_bt <- read_csv('hourly-traffic-on-metropolitan-transportation-authority-mta-bridges-and-tunnels-beginning-2010.csv') # MTA bridges & tunnels data
MTA_fares <- read_csv('fare-card-history-for-metropolitan-transportation-authority-mta-beginning-2010.csv')              # MTA subway fare data  
MTA_KPI <- read_csv('metropolitan-transportation-authority-mta-performance-indicators-per-agency-beginning-2008.csv')    # MTA KPI performance data 
NYPD <- read_csv('nypd-motor-vehicle-collisions.csv', col_types = list(DATE = col_character()))                          # NYPD collision data 
```


I was inspired by this [CityLab](https://www.citylab.com/perspective/2018/12/new-york-city-electric-bikes-transit-crisis-de-blasio/577969/) article detailing the impending transportation crisis due to the MTA L Line shutdown. Specifically, I am interested in observing whether the shutdown contributes to an increase in cyclist injuries due to motor vehicle collisions as about [275,000 L Line riders](https://ny.curbed.com/2018/10/3/17925038/nyc-subway-l-train-shutdown-mitigation-alternatives) seek alternatives.  
  
[NYC DOT](http://web.mta.info/sandy/CanarsieTunnelRebuildingProcess.html#sdtr) has predicted about 2% of L Line riders to switch to cycling, while about 79% are predicted to move to other subway lines. I could investigate the dispersion of L Train riders to other options as a whole if data is available, but for now I will focus on cyclist safety as it's more approachable with currently available data.  

Let's start by looking at some performance metrics for the MTA subway network as a whole.  

```{r echo=FALSE}
# convert Period to date variable
transit <- MTA_KPI

transit$Period <- paste(transit$Period, '-01', sep = '')

transit$Period <- transit$Period %>%
  ymd()

 # filter observations to only NYC Transit (Subway and Bus)
transit <- transit %>% 
  dplyr::filter(MTA_KPI$`Agency Name` == 'NYC Transit') %>%
  group_by(Period) 
```

```{r echo=FALSE, message=FALSE}
# plot ridership 
plt.NYC.riders <- transit %>% 
  dplyr::filter(`Indicator Name` == 'Total Ridership - Subways') %>%
  ggplot(aes(x = Period, y = `Monthly Actual`)) + 
  geom_point(color = 'steelblue4', alpha = .7) + 
  geom_smooth(method = 'lm') + 
  ylab('Total Monthly Ridership') +
  scale_y_continuous(labels = comma) +
  theme_classic()

# convert to plotly object
ggplotly(plt.NYC.riders) %>% 
  layout(title = 'Total Monthly Ridership - NYC Transit (All Lines)',
  titlefont = list(size = 17)) %>% 
  config(displayModeBar = FALSE)

```

```{r echo=FALSE, message=FALSE}
# plot On-Time Performance for all subway lines
plt.NYC.OTP <- transit %>% 
  dplyr::filter(`Indicator Name` == 'On-Time Performance (Terminal)') %>%
  ggplot(aes(x = Period, y = `Monthly Actual`)) + 
  geom_point(color = 'steelblue4', alpha = .7) + 
  geom_smooth(method = 'lm') + 
  ylab('Monthly On-Time Performance') +
  theme_classic()

# convert to plotly object
ggplotly(plt.NYC.OTP) %>% 
  layout(title = 'On-Time Performance - NYC Transit (All Lines)',
  titlefont = list(size = 17)) %>% 
  config(displayModeBar = FALSE)
```

```{r echo=FALSE, message=FALSE}
# plot subway wait assessment for all subway lines
plt.NYC.SWA <- transit %>% 
  dplyr::filter(`Indicator Name` == 'Subway Wait Assessment') %>%
  ggplot(aes(x = Period, y = `Monthly Actual`)) + 
  geom_point(color = 'steelblue4', alpha = .7) + 
  geom_smooth(method = 'lm') +
  ylab('Subway Wait Assessment') +
  theme_classic()

# convert to plotly object
ggplotly(plt.NYC.SWA) %>% 
  layout(title = 'Subway Wait Assessment - NYC Transit (All Lines)',
  titlefont = list(size = 17)) %>% 
  config(displayModeBar = FALSE)

```

```{r echo=FALSE, message=FALSE}
# plot mean distance between failures for all subway lines 
plt.NYC.fail <- transit %>% 
  dplyr::filter(`Indicator Name` == 'Mean Distance Between Failures - Subways') %>%
  ggplot(aes(x = Period, y = `Monthly Actual`)) + 
  geom_point(color = 'steelblue4', alpha = .7) + 
  geom_smooth(method = 'lm') +
  ylab('Mean Distance Between Failures') +
  scale_y_continuous(labels = comma) +
  theme_classic()

# convert to plotly object
ggplotly(plt.NYC.fail) %>% 
  layout(title = 'Mean Distance Between Failures - NYC Transit (All Lines)',
  titlefont = list(size = 17)) %>% 
  config(displayModeBar = FALSE)

```

Next, the L Line specifically: 

```{r echo=FALSE, message=FALSE}
# plot On-Time Performance for the L Line 
plt.L.OTP <- transit %>% 
  dplyr::filter(`Indicator Name` == 'OTP (Terminal) - L Line') %>% 
  ggplot(aes(x = Period, y = `Monthly Actual`)) + 
  geom_point(color = 'steelblue4', alpha = .7) + 
  geom_smooth(method = 'lm') +
  ylab('Monthly On-Time Performance') + 
  theme_classic()  

# convert to plotly object
ggplotly(plt.L.OTP) %>% 
  layout(title = 'L Line On-Time Performance has decreased, but is relatively high', 
  titlefont = list(size = 17)) %>% 
  config(displayModeBar = FALSE)

```

Now let's look at some two-wheeled data.
<br>

```{r echo=FALSE}
# convert date to a date variable
NYPD$DATE <- NYPD$DATE %>% 
  substr(start = 1, stop = 10) %>% 
  ymd()

# create summarized subset of NYPD data 
sub <- NYPD 

# summarize by week
sub <- sub %>% 
  filter(DATE > 16438) %>% 
  mutate(week = floor_date(DATE, 'weeks', week_start = 7)) %>% 
  group_by(week) %>% 
  summarize(cyclist_injuries = sum(`NUMBER OF CYCLIST INJURED`))

```

```{r echo=FALSE, message=FALSE}
# plot cyclist injuries per week 
plt.sub <- sub %>% 
  ggplot(aes(x = week, y = cyclist_injuries)) +
  geom_point(color = 'steelblue4', alpha = .7) + 
  geom_smooth(method='lm') + 
  xlab('Week') +
  ylab('Cyclist Injuries') + 
  theme_classic()

# convert to plotly object
ggplotly(plt.sub) %>% 
  layout(title = 'Weekly Cyclist Injuries Resulting from Motor Vehicle Collisions',
  titlefont = list(size = 17)) %>% 
  config(displayModeBar = FALSE)
```

```{r}
# convert weekly crash data to a time-series object
ts_injuries <- ts(sub$cyclist_injuries, frequency = 52)
# plot autocorrelation function
acf(ts_injuries)
```

```{r echo=FALSE}
arima1 <- auto.arima(ts_injuries, stepwise = FALSE, trace = FALSE)
summary(arima1)
```


Planned Updates:  
lines of questioning - EDA  
how many riders of L train  
how many cyclists  
how many bus riders  
refer to article predictions of where capacity will move to   
present general transit trends maybe  
show map of cyclist accidents?  
show time-series trends of accidents  
conclude with whether L train has increased accidents  
leaflet timeline is difficult at best, consider gganimate  
gganimate issues installing with R 3.5.1  
ARIMA and intervention analysis/Interrupted Time Series Analysis    
consider using citibike data as a proxy for bike usage   

***

Radix is a publication format for scientific and technical writing, native to the web. 

Learn more about using Radix at <https://rstudio.github.io/radix>.





